---
title: "Casitas-scratch-paper"
format: html
editor: visual
---

```{r, libraries}
# model packages
library(MASS) # have to read this in before tidyverse 
library(lme4)
library(glmmTMB) # ok if you don't have this - just comment it out

# diagnostics and model info
library(DHARMa)
library(MuMIn)
library(ggeffects)
library(lmtest)
library(broom)

library(here)
library(naniar)
library(tidyverse)
library(GGally)
library(tidyr)
library(corrplot)
library(dplyr)
library(janitor)
```

# Scratch paper

## Problem two

Questions: How does seed count vary with plot type (shrub or open), plant species, and total number of inflorescences? Is there a simpler model that explains seed count, and if so, what is it?

Independent Variable: Seed Count

Predictor Variable: Plot type (shrub or open) [binary], plant species , total number of inflorescences

```{r}

#Load in Problem 2 data
pollinator <- here("data", "shrubstudy_pollinator.ms.data.csv") |> 
  read.csv()

seed_ctwt <- here("data", "shrubstudy_seed_ctwt.ms.data.csv") |> 
  read.csv()

individual_flower <- here("data", "shrubstudy_individual_flower_counts.ms.data.csv") |> 
  read.csv()

community_flower <- here("data", "shrubstudy_community_flower_counts.ms.data.csv") |> 
  read.csv()
```

```{r}
View(pollinator)
```

```{r}
View(seed_ctwt)
colnames(seed_ctwt)
```

```{r}
View(individual_flower)
```

```{r}
View(community_flower)
```

### Restart 

Idk, had a problem with the fucntion select 

```{r}
seed_ctwt <- here("data", "shrubstudy_seed_ctwt.ms.data.csv") |> 
  read.csv()
```

```{r}
gg_miss_var(seed_ctwt)
```

```{r}
# from peppa 
#Clean data


#seed_ctwt_subset <- seed_ctwt |> 
  #select(c("treatment", "species", "total_nr_infl", "nr_seeds")) |> #Plot type, plant species, total number of inflorescences - select relevant columns 
  #drop_na(nr_seeds) #Thats a lot of data missing... should we do anything? 

#Not quite sure what the total number of inflorescences vs inflorescence collected means
```

```{r}
seed_clean <- seed_ctwt %>% 
  clean_names()
```

```{r}
colnames(seed_clean)
View(seed_clean)
```

```{r}
seed_ctwt_subset2 <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds) %>% 
  drop_na(nr_seeds)

```

```{r}
View(seed_ctwt_subset2)
```

```{r}
# linear model, we know this is wrong
seed_L1 <- lm(nr_seeds ~ treatment + species + total_nr_infl, data = seed_ctwt_subset2)

# generalized linear model with Poisson distribution
seedctwt2 <-glm(nr_seeds ~ treatment + species + total_nr_infl, data = seed_ctwt_subset2, family = "poisson")
seedctwt2.a <- glm(nr_seeds ~ treatment + species + total_nr_infl, data = seed_ctwt_subset2, family = "poisson")

# generalized linear model with negative binomial distribution
seedctwt3 <- glm.nb(nr_seeds ~ treatment + species + total_nr_infl, data = seed_ctwt_subset2)
seedctwt3.a <- glmmTMB(nr_seeds ~ treatment + species + total_nr_infl, data = seed_ctwt_subset2, family = "nbinom2")

```

```{r}
plot(simulateResiduals(seed_L1))
```

```{r}
plot(simulateResiduals(seedctwt2))
```

```{r}
plot(simulateResiduals(seedctwt2.a))
```

```{r}
plot(simulateResiduals(seedctwt3))
```

```{r}
plot(simulateResiduals(seedctwt3.a))
```

### take-0.5. 

```{r}

```



### take-1, [aspect]
to much multicollinerarity
```{r}
# trying aspect
seed_ctwt_subset_as <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, aspect) %>% 
  drop_na(nr_seeds,aspect)
```

```{r}
seedctwtas <- glmer(nr_seeds ~ treatment + species + (1|aspect) + total_nr_infl, data = seed_ctwt_subset_as, family = "poisson" )
#Warning message:
#In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  #Model is nearly unidentifiable: very large eigenvalue
 #- Rescale variables?
```

```{r}
plot(simulateResiduals(seedctwtas))
```

### take-2, [nr_infl_coll], got it to work but fail 
```{r}
# trying nr_infl_coll
seed_ctwt_subset_nrf <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds,nr_infl_coll) %>% 
  drop_na(nr_seeds,nr_infl_coll)
```

```{r}
seedctwtnrf <- glmer(nr_seeds ~ treatment + species + (1|nr_infl_coll) + total_nr_infl, data = seed_ctwt_subset_nrf, family = "poisson" )
```

```{r}
seedctwtnrf_5 <- glmer.nb(nr_seeds ~ treatment + species + (1|nr_infl_coll) + total_nr_infl, data = seed_ctwt_subset_nrf)
```

```{r}
plot(simulateResiduals(seedctwtnrf_5))
```

```{r}
plot(simulateResiduals(seedctwtnrf))
```


### take-3, [wt_seeds_mg], fail 
```{r}
# trying wt_seeds_mg
seed_ctwt_subset_wtsm <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, wt_seeds_mg) %>% 
  drop_na(nr_seeds, wt_seeds_mg)
```

```{r}
seedctwtwtsm <- glmer(nr_seeds ~ treatment + species + (1|wt_seeds_mg) + total_nr_infl, data = seed_ctwt_subset_wtsm, family = "poisson" )
#Warning message:
#In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  #Model failed to converge with max|grad| = 0.0199853 (tol = 0.002, component 1)
```

```{r}
plot(simulateResiduals(seedctwtwtsm))
```

```{r}
seedctwtwtsm_5 <- glmer.nb(nr_seeds ~ treatment + species + (1|wt_seeds_mg) + total_nr_infl, data = seed_ctwt_subset_wtsm)
```

```{r}
plot(simulateResiduals(seedctwtwtsm_5))
```


### take-4 , [dist_fr_shr], fail
```{r}
# trying dist_fr_shr
seed_ctwt_subset_dfs <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, dist_fr_shr) %>% 
  drop_na(nr_seeds, dist_fr_shr)
```

```{r}
#seedctwtdfs <- glmer(nr_seeds ~ treatment + species + (1|dist_fr_shr) + total_nr_infl, data = seed_ctwt_subset_dfs, family = "poisson" )
#Error in `contrasts<-`(`*tmp*`, value = contr.funs[1 + isOF[nn]]) : 
  #contrasts can be applied only to factors with 2 or more levels
```

```{r}
#plot(simulateResiduals(seedctwtdfs))
```

### take-5 [shrub_num]
to much multicollinearity 

```{r}
seed_ctwt_subset_sn <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, shrub_num) %>% 
  drop_na(nr_seeds, shrub_num)
```

```{r}
seedctwtwtsn <- glmer(nr_seeds ~ treatment + species + (1|shrub_num) + total_nr_infl, data = seed_ctwt_subset_sn, family = "poisson" )
#Warning message:
#In checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
  #Model is nearly unidentifiable: very large eigenvalue
# - Rescale variables?
```

```{r}
plot(simulateResiduals(seedctwtwtsn))
```

### take-6 [plant_nr]
to much multicollinearity 
```{r}
seed_ctwt_subset_pn <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, plant_nr) %>% 
  drop_na(nr_seeds, plant_nr)
```

```{r}
View(seed_ctwt_subset_pn)
colnames(seed_ctwt_subset_pn)
```

```{r}
seedctwtwtpn <- glmer(nr_seeds ~ treatment + species + (1|plant_nr) + total_nr_infl, data = seed_ctwt_subset_pn, family = "poisson" )
```

```{r}
plot(simulateResiduals(seedctwtwtpn))
```


### take-7 [tag_num]
to much multicollinearity 
```{r}
seed_ctwt_subset_tgnm <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, tag_num) %>% 
  drop_na(nr_seeds, tag_num)
```

```{r}
seedctwtwttgnm <- glmer(nr_seeds ~ treatment + species + (1|tag_num) + total_nr_infl, data = seed_ctwt_subset_tgnm, family = "poisson" )
```

```{r}
plot(simulateResiduals(seedctwtwttgnm))
```

### take-8
to much multicollinearity 
```{r}
seed_ctwt_subset_dc <- seed_clean %>% 
  select(treatment, species, total_nr_infl, nr_seeds, date_collected) %>% 
  drop_na(nr_seeds, date_collected)
```

```{r}
seedctwtwtdc <- glmer(nr_seeds ~ treatment + species + (1|date_collected) + total_nr_infl, data = seed_ctwt_subset_dc, family = "poisson" )
```

```{r}
plot(simulateResiduals(seedctwtwtdc))
```





## Trying to merge the datasets now 








